---
title: "Day 13 Lab"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Day 13 Lab - Introduction to R

The purpose of this lab is to give you a *basic* introduction to R
syntax.

Getting help:

If you know what function you want to use but need a bit more
information try `help(function)` or `?function`.

If you don't know what function you want to use, but know a keyword
associated with it (for example you know you want to find the variance
but don't know what function to use, try `help.search(variance)` or
`??variance`

## Section 1 - Basic Data Types:

#### (1) Practice creating variables in R. Make a few that are numeric, some logical and a few character type objects. Use the `class` and `typeof` functions to verify that you have done it.

```{r}
x <- "hello"
y <- 5L
z <- 3.14
a <- TRUE
b <- 4i
class(x)
class(y)
class(z)
class(a)
class(b)
```

#### (2) What are the arithmetic and logical operators in R? Practice using them here.

```{r problem2}
add <- 5 + 6
sub <- 50 - 43
multi <- 4 * 33
div <- 8 / 2

```

#### (3) Another way to check the data type is the `is` function. With this function you can ask if an object is of a certain class and get a logical value back out:

-   `is.numeric(5)`
-   `is.character("a")`

Try these out using different object types.

```{r problem3}
is.numeric(5)
is.character("a")
```

#### (4) A complementary function is `as`. This works the same way as `is` but will instead try to convert between object types. Try it out using different object types.

*Example: `as.numeric("5")`*

```{r problem4}
as.numeric("5")
```

#### (5) In R the `length` function will tell you how long your object is and the `nchar` function will tell you how many characters are in a string. Try them both out with a character object. What is the difference?

```{r problem5}

```

#### (6) Grabbing a portion of a character object in R is a more difficult than in python. We cannot use indexing the way we did in python. Now we are going to rely on the `substr` function. This will allow you to take a segment of a string. Use the string `"R is awesome!"` and try to select the word `"awesome"`.

`substr(aString, start, end)` *Note: End is inclusive*

```{r problem6}
string = "R is awesome!"
substr(string, 6, 12)
```

#### (7) Indexing in python starts at 0. What value does indexing in R start with? Does negative indexing work in R?

```{r problem7}
#R indexing starts at 1 rather than 0
list = 1:5
list[-1]
```

## Section 2 - Vectors

<font color = #800080><b>Vectors</b></font> are series of **components**
that all have the same type.

#### (8) Use the `c`, `seq`, and `rep` function to make a few vectors. Store those objects in variables.

```{r problem8}
vec1 = c("hi", "this", "is", "a", "vector")
vec2 = seq(1:10)
vec3 = rep(3, 4)
vec4 = seq(11:20)
```

#### (9) What are the classes of the vectors you created? What are their lengths?

```{r problem9}
class(vec1)
class(vec2)
class(vec3)
length(vec1)
length(vec2)
length(vec3)
```

#### (10) What operators work with vectors?

```{r problem10}
vec5 = vec2 + vec4
print(vec5)
```

#### (11) You can use indexing with the vector objects. Pracitce selecting single componets.

```{r problem11}
vec5[2]
```

#### (12) You can select more than one component using another vector. See the example below and practice using this feature. Look up the `seq` function and make sure you know how it works.

```{r problem12}

ex_vec = c('Vector', 'of', 'character', 'components')
ex_vec[seq(1,4,2)]

```

#### (13) Arithmetic vectors will apply to an entire vector. Create a vector of numbers from 1 to 15 and divide the vector by 2.

```{r problem13}
numbers = seq(1:15)/2
print(numbers)
```

#### (14) R has a useful function called `which`. This will take in a logical vector and return to you the index locations for the TRUE values. Use the `which` function to locate the values that are less than 5 from your random numbers. Then, use the `which` function to select all values from your vector that are less than 5.

```{r problem14}
logs = numbers[which(numbers < 5)]
print(logs)
```

#### (15) R has many functions built-in for generating random numbers. Use `rnorm` to generate a vector of length 50 from a normal distribution with mean 10 and sd of 5.

`rnorm(n, mean, sd)`

\*Note: R has many built in distributions such as uniform
`runif(n, min, max)`, poisson `rpois(n, lambda)`, binomial
`rbinom(n, size, prob)` and many more `?Distributions`

```{r problem15}
randoms <- rnorm(50, 10, 5)
print(randoms)
```

#### (16) What is the minimum and maximum of your vector of random numbers?

*Hint: min and max work just like they do in python*

```{r problem16}
print(max(randoms))
print(min(randoms))
```

#### (17) `which.min` and `which.max` will tell you the location of your min and max. Use those now.

```{r problem17}
print(which.min(randoms))
print(which.max(randoms))
```

#### (18) Components in vectors can be given names. The `names` function will allow you to view the names of your components as well as set the names. Create a new vector and give each component a name.

*I have provided some code as an example.*

```{r problem18}
my_name = c(fname = "Lauren", lname = "Jepsen")
print(my_name)
print(names(my_name))

numbers = 1:3
names(numbers) = c("one", "two", "three")
print(numbers)

# Put your code here:
my_name_2 <- c(fname = "Aaron", lname = "Standefer")
print(my_name_2)
print(names(my_name_2))

nums = 4:6
names(nums) = c("four", "five", "six")
print(nums)
```

## Section 3 - Matrices

These are 2D matrices the way we would think of them in math.

They can be created using the `rbind`, `cbind`, `matrix`, and
`as.matrix` functions.

#### (19) Run the following code then look at the resulting variables. What are the differences between mat1 and mat2?

```{r problem19}
mat1 = matrix(1:10, nrow = 5, ncol = 2, byrow = T)
mat2 = matrix(1:10, nrow = 5, ncol = 2, byrow = F)
print(mat1)
print(mat2)
#I changed one of the byrow variables to False to illustrate what i think
#you were trying to show.  Major difference is how the matrix is filed.
```

#### (20) Create two vectors of equal length. Use `rbind` and `cbind` on them. What are the differences between `rbind` and `cbind`?

*Example: `rbind(1:5,6:10)`*

```{r problem20}

```

#### (21) The `apply` funciton will let you perofrm operations using entire rows or colums. Use `apply` to get row and column sums from mat1.

`apply(X, MARGIN, FUN, ..., simplify = TRUE)` \* MARGIN = 1 for rows \*
MARGIN = 2 for colums

```{r problem21}

```

#### (22) Just like with vectors, you can provide names for your rows and columns. Now use the `rownames` and `colnames` function to change the names associated with one of your matrices.

```{r problem22}

```

#### (23) Practice taking pieces of your matrix. What are the different ways this can be done?

```{r problem23}

```

#### (24) `dim`, `nrow`, and `ncol` are useful functions to check the size of your matrix. Try them now, how do they work? Try also the `length` function.

```{r problem24}

```

## Section 4 - Data Frames

Data frames are lists of vectors of the same length. It allows you to
combine values of different type into a matrix like format.

These are most often used for data tables and can be compared to
Python's pandas data frame.

#### (25) Use the following to create a sample data frame. Look at your data frame.

```{r problem25}
num = c(1,2,3)
let = c("a", "b", "c")
logic = c(T, F, T)

df = data.frame(num, let, logic)

```

#### (26) What are the column names of your data frame?

*The function `colnames` will tell you what your column names are.*

```{r problem26}

```

#### (27) You can use `$` notation to access entire columns. Try

`df$num`. What does that give you?

```{r problem27}

```

#### (28) You can also access information in other ways. Try getting different pieces of information back out of your data frame

`df[rnum, colnum]` `df[,"colnumname"]`

```{r problem28}

```

#### (29) Create your own data frame and practice accessing different values in your data frame.

```{r problem29}

```

#### (30) R has loops but they are inefficient. Use a loop to pring every element inside the first column of your data frame.

*Your for loop will look something like this:*

for (loopVariable in Sequence){ print(loopVariable) }

```{r problem30}

```

## Section 5 - Lists

A list is a generic vector that can contain objects of different types.

Lists can have named components.

This is often the object of choice for output from various functions.

#### (31) Try creating the following list. Look at it.

```{r problem31}
n = c(2, 3, 5)
s = c("aa", "bb", "cc", "dd", "ee")
b = c(TRUE, FALSE, TRUE, FALSE, FALSE)
x = list(n,s,b, 3)
```

#### (32) Does your list have names? If it does not, name your components `n`, `s`, `b`.

```{r problem32}

```

#### (33) There are many ways to access vectors inside of lists. The following represent 3 different ways to access the vector `n`. Run them each and then try to access the vector `b`.

*Many of the objects we will work with over the next couple of days will
be lists. Know how to access their components.*

```{r problem33}
x$n # Works with named components
x[['n']] # Double square brackets for accessing elements of a list
x[[1]]
```

#### (34) Just like we had `apply` for array objects we have `lapply` for lists. Use lapply on x to find the components that are numeric.

`lapply(aList, FUN)`

```{r problem34}

```

#### (35) What does the `attributes` function tell you about your list?

*x doesn't have a lot of information associated with it. This will be
more useful when you have outputs from functions*

```{r problem35}

```

## Section 6 - Plotting

R has some really great plotting features. We are going to going to very
briefly explore some of the most basic plotting features here.

#### (36) Generate a vector of normally distributed random numbers. Pick your mean and sd.

```{r problem36}

```

#### (37) What is the mean and standard deviation of your vector? How closely do they match the mean and sd you used in question 36?

*Use the mean and sd functions.*

```{r problem37}

```

#### (38) Use the plot function to plot your values. Look up at the help documentation and see if you can't figure out how to add a y-axis label and a title.

```{r problem38}

```

#### (39) The above was a terrible representation of your data. Try generating a histogram `hist` and a `boxplot` of your numbers.

*Try to make these plots look nice. Change colors, add labels, etc.*

```{r problem39}

```

## Section 7 - Functions

Just like with other programming languages, you can define your own
functions.

#### (40) Create a function that will slice a string in a way that is similar to python. We want this function to take in 4 parameters: string, start=1, stop = nchar(string), step=1.

Here are some building blocks to help: \* `strsplit(string, "")` will
create a list containing a vector. The components of the vector will
have individual characters. \* `strsplit(string, "")[[1]]` pulls the
vector out of the list. This vector will have each individual character
in it. \* `strsplit(string, "")[[1]][inds]` This will grab the
characters at the specified index values `inds`. `inds` can be numeric
or logical. \* `paste(***, collapse = "")` can collapse the vector of
characters down into a single string. Replace `***` with
`strsplit(string, "")[[1]][inds]`

A function can be created as follows:

functionName = function(parameter1, parameter2 = defaultValue){ code \#
automatically returns last value produced.}

```{r problem40}

```
