---
title: "Homework5"
output: html_document
---

# Homework 4 and 5

The purpose of this assignment is to evaluate the structural similarity between human HBB and Cytoglobin. As this assignment is both longer and harder than others we have done in the course I am breaking it into two parts. The first part is due Monday and the second part will be due Tuesday Morning.

### Homework 5:
Submit your R code for this assignment as well.

#### (1) Load your HBB and Cytoglobin proteins you choose in Homework 4 into R. What chain from the PDB files are associated with your proteins?

```{r}

```

#### (2) Construct a sequence alignment between HBB and Cytoglobin. Evaluate different scoring schemes and pick one that will give you the optimal alignment between the two. Comment on your scoring scheme and how changing scoring parameters altered your alignments. What scoring scheme did you end up using?

In this case I am using the term **optimal alignment** to refer to a scoring scheme that leads to the highest sequence identity between the two proteins.

*Note: Making your match scores very high will inflate your score. We want to compare identity here.* 

```{r}

```

#### (3) What is the sequence identity and sequence similarity between the two proteins? Comment on your findings.

```{r}

```

#### (4) Use `aln2html` to view your alignment. Upload that file with your assignment.

```{r}

```

#### (5) What is the RMSD between your two proteins?

```{r}

```

#### (6) Load both structures into PDB's molecule viewer and generate an image. Comment on the (dis)similarity between the two structures. Include that image with your homework submission. 

#### (7) Write a short paragraph summarizing what you learned over the last two homeworks.